R and Python clients for GBIF species occurrence data

### 2016-08-05 manuscript.tex
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

      \address[cstar]{University of California, Berkeley, CA, USA}    
    \address[boettig]{University of California, Berkeley, CA, USA}

  \begin{abstract}
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \item
    xxx
  \item
    xxx
  \item
    xxx
  \item
    xxxx
  \end{enumerate}
  \end{abstract}

@@

@@

\subsection{The rgbif package}\label{the-rgbif-package}

@@

\subsection{The pygbif library}\label{the-pygbif-library}

\texttt{pygbif} (Chamberlain) is a Python library for working with GBIF
data in the Python programming environment ({\textbf{???}}). Python is a
general purpose programming language used widely in all sectors, and for
all parts of software development including server and client side use
cases. Python is used exclusively in some scientific disciplines (e.g.,
astronomy), and has partial usage in other disciplines. A Python client
for GBIF data is an important tool given the even wider usage of Python
than R, though maybe slightly less than R for ecology/biology.

@@

The \texttt{pygbif} library is less mature and complete than the R
package. It also uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{pygbif} is developed publicly on GitHub at
\href{https://github.com/sckott/pygbif}{\url{https://github.com/sckott/pygbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{pygbif} can be
installed from \href{https://pypi.python.org/pypi/pygbif}{pypi}, the
distribution network for Python libraries.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2737).

@@

The GBIF backbone taxonomy is used in GBIF to have a consistent way to
refer to taxonomic names throughout their services. The backbone has
4410899 unique names and 2497114 species names. The backbone taxonomy is
also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# well known text}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\CommentTok{# bounding box}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{125.0}\NormalTok{,}\FloatTok{38.4}\NormalTok{,-}\FloatTok{121.8}\NormalTok{,}\FloatTok{40.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"dismo"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"maptools"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn$data$nubKey))}
\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
\NormalTok{df <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{class}\NormalTok{(x$data)) %in%}\StringTok{ "data.frame"} \NormalTok{]}
\NormalTok{df <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
\NormalTok{df2 <-}\StringTok{ }\NormalTok{df[,}\KeywordTok{c}\NormalTok{(}\StringTok{'decimalLongitude'}\NormalTok{,}\StringTok{'decimalLatitude'}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"RCurl"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Bounding lat/long data is from
\href{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}{here}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{getURL}\NormalTok{(}\StringTok{'https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt'}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}\DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{sep=}\StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{))}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{dat$city, }\DataTypeTok{minlon=}\NormalTok{dat$minlon, }\DataTypeTok{minlat=}\NormalTok{dat$minlat, }\DataTypeTok{maxlon=}\NormalTok{dat$maxlon, }\DataTypeTok{maxlat=}\NormalTok{dat$maxlat)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{getdata <-}\StringTok{ }\NormalTok{function(x)\{}
  \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords)$meta$count}
  \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{x[}\StringTok{'city'}\NormalTok{], }\DataTypeTok{richness=}\NormalTok{num, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by=}\StringTok{"city"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat+maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon+maxlon)/}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

From
\href{http://esp.cr.usgs.gov/data/little/}{\url{http://esp.cr.usgs.gov/data/little/}}.
And save shapefile in same directory

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{), }\DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
\CommentTok{# add shape from Atlas of US Trees}
\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# add Gbif presence points}
\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{, }\StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{, }\DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
interfaces to GBIF's application programming interface (API) - a
powerful tool for making research using species occurrence data
reproducible. In fact, the \texttt{rgbif} package has already been used
in 22 scholarly publications (as of 2015-11-14).

@@

The \texttt{pygbif} library is still in early development, and will
greatly benefit from any feedback and use cases.

One area of focus in the future is to attempt to solve many use cases
that have been brought up with respect to GBIF data. For example, some
specimens are included in GBIF that are located in botanical gardens.
For many research questions, researchers are interested in ``wild'' type
occurrences, not those in human curated scenarios. Making removal of
these occurrences easy would be very useful, but is actually quite a
hard problem.

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
\href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
Software can be found at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
under an MIT license.

@@

Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{Pygbif: Interface to
the global 'biodiversity' information facility 'aPI'}.

Chamberlain S. \emph{Rgbif: Python client for gBIF}.

>>>>

      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
CA, USA}
    \address[boettig]{rOpenSci, Department of Enivornmental Science, Policy and Management,
University of California, Berkeley, CA, USA}

  \begin{abstract}
  Corresponding Author:

  Scott Chamberlain

  rOpenSci, Museum of Paleontology, University of California, Berkeley,
  CA, USA

  Email address:
  \href{mailto:scott@ropensci.org}{\href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}}

  \newpage

  Background. The number of individuals of each species in a given
  location forms the basis for many sub-fields of ecology and evolution.
  Data on individuals, including which species, and where they're found,
  can be used for a large number of research questions. Global
  Biodiversity Information Facility (hereafter, GBIF) is the largest of
  these. Programmatic clients for GBIF would make research dealing with
  GBIF data much easier.

  Methods. We have developed a client to access GBIF data for each of the
  R and Python programming languages: \texttt{rgbif} and \texttt{pygbif}.

  Results. For both clients, we describe their design and utility, and
  demonstrate many use cases.

  Discussion. Programmatic acccess to GBIF will facilitate more open and
  reproducible science. We hope that all scientists using GBIF data will
  do so with one of these two clients.
  \end{abstract}

@@

\newpage

@@

\subsection{rgbif}\label{rgbif}

@@

\subsection{pygbif}\label{pygbif}

\texttt{pygbif} (Chamberlain) is a Python library for working with GBIF
data in the Python programming environment. Python is a general purpose
programming language used widely in all sectors, and for all parts of
software development including server and client side use cases. Python
is used exclusively in some scientific disciplines (e.g., astronomy),
and has partial usage in other disciplines. A Python client for GBIF
data is an important tool given the even wider usage of Python than R,
though maybe slightly less than R for ecology/biology disciplines.

@@

The \texttt{pygbif} library is less mature and complete than the R
package. It also uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{pygbif} is developed publicly on GitHub at
\url{https://github.com/sckott/pygbif}, where development versions of
the package can be installed, and bugs and feature requests reported.
Stable versions of \texttt{pygbif} can be installed from
\href{https://pypi.python.org/pypi/pygbif}{pypi}, the distribution
network for Python libraries.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (16057).

@@

The GBIF backbone taxonomy is used in GBIF to have a consistent way to
refer to taxonomic names throughout their services. The backbone has
5194833 unique names and 2420842 species names. The backbone taxonomy is
also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).

@@

Spatial search, based on well known text format (Herring, 2011), or a
bounding box set of four coordinates. The well known text string and the
bounding box in the below example specify the same rectangular area in
California, centering approximately on Sacramento. Whereas the bounding
box format requires
\texttt{longitude SW corner, latitude SW corner, longitude NE corner, latitude NE corner},
the well known text string requires an extra long/lat pair to close the
polygon.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# well known text}
\NormalTok{wkt <-}\StringTok{ 'POLYGON((-122.6 39.9,-120.0 39.9,-120.0 37.9,-122.6 37.9,-122.6 39.9))'}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\CommentTok{# bounding box}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,-}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Get only occurrences with lat/long data using the \texttt{hasCoordinate}
parameter

@@

There's no equivalent interface in \texttt{pygbif} yet.

@@

\section{R vs.~Python}\label{r-vs.python}

Both R and Python are commonly used in science, and can be used for
similar tasks. Python, however, is a more general programming lanugae,
and can be used in more contexts than R can be used in.

The \texttt{rgbif} and \texttt{pygbif} clients can do a lot of the same
tasks. We envision \texttt{rgbif} as being more common in workflows of
academics asking research questions, whereas \texttt{pygbif} can do that
as well, but may be more easily used in a website.

The R client \texttt{rgbif} has much more development time than
\texttt{pygbif}, but through time \texttt{pygbif} will become equally
mature.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"dismo"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"maptools"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn$data$nubKey))}
\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
\NormalTok{df_data <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x$data) %in%}\StringTok{ "tbl_df"}\NormalTok{)) ]}
\NormalTok{df_data <-}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
\NormalTok{df2 <-}\StringTok{ }\NormalTok{df_data %>%}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"httr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Bounding lat/long data is from
\url{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'https://raw.githubusercontent.com/amyxzhang/}
\StringTok{boundingbox-cities/master/boundbox.txt'}
\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
  \DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
  \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{city =} \NormalTok{dat$city, }\DataTypeTok{minlon =} \NormalTok{dat$minlon, }
  \DataTypeTok{minlat =} \NormalTok{dat$minlat, }\DataTypeTok{maxlon =} \NormalTok{dat$maxlon, }
  \DataTypeTok{maxlat =} \NormalTok{dat$maxlat,}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

A helper function to get count data. GBIF has a count API, but we can't
use that with a geometry search as that API doesn't support geospatial
search. We can however use the search API via \texttt{occ\_search()} and
set \texttt{limit = 1} so that we

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{getdata <-}\StringTok{ }\NormalTok{function(x)\{}
  \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)$meta$count}
  \KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{city =} \NormalTok{x[}\StringTok{'city'}\NormalTok{], }
    \DataTypeTok{richness =} \NormalTok{num, }
    \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
  \NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by =} \StringTok{"city"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat +}\StringTok{ }\NormalTok{maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon +}\StringTok{ }\NormalTok{maxlon)/}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

From \url{http://esp.cr.usgs.gov/data/little}. And save shapefile in
same directory

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
     \DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{,}
     \DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
\CommentTok{# add shape from Atlas of US Trees}
\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{),}
     \DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# add Gbif presence points}
\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude,}
       \DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
       \DataTypeTok{legend =} \StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{,}
       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
interfaces to GBIF's application programming interface (API) - a
powerful tool for working with species occurrence data, and facilitating
reproducible research. In fact, the \texttt{rgbif} package has already
been used in 22 scholarly publications (as of 2015-11-14).

@@

The \texttt{pygbif} library is in early development, and will greatly
benefit from any feedback and use cases.

One area of focus in the future is to attempt to solve many use cases
that have been brought up with respect to GBIF data. For example, some
specimens are included in GBIF that are located in botanical gardens.
For many research questions, researchers are interested in ``wild'' type
occurrences, not those in human curated scenarios. Making removal of
these occurrences easy would be very useful, but is actually quite a
hard problem. There are many other problems like this.

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
\url{https://github.com/sckott/msrgbif}. Software can be found at
\url{https://github.com/ropensci/rgbif} and
\url{https://github.com/sckott/pygbif}, both under MIT licenses. We
thank all the users that have used \texttt{rgbif} and \texttt{pygbif}
and have given feedback and reported bugs. In addition, we greatly
appreciate the all contributions to the two packages, found at
\url{https://github.com/ropensci/rgbif/graphs/contributors} and
\url{https://github.com/sckott/pygbif/graphs/contributors}.

@@

Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{rgbif: An r interface
to the global 'biodiversity' information facility API}.

Chamberlain S. \emph{pygbif: A python interface to the global
biodiversity information facility API}.

### 2016-08-05 use-cases/city-diversity.Rmd
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

```{r load}
library("rgbif")
library("ggplot2")
library("plyr")
library("RCurl")
library("RColorBrewer")
```

@@

Bounding lat/long data is from [here](https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt).

```{r box}
rawdat <- getURL('https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt')
dat <- read.table(text = rawdat, header = FALSE, sep="\t", col.names=c("city","minlat","maxlon","maxlat","minlon"))
dat <- data.frame(city=dat$city, minlon=dat$minlon, minlat=dat$minlat, maxlon=dat$maxlon, maxlat=dat$maxlat)
```

```{r newsearch}
getdata <- function(x){
  coords <- as.numeric(x[c('minlon','minlat','maxlon','maxlat')])
  num <- occ_search(geometry = coords)$meta$count
  data.frame(city=x['city'], richness=num, stringsAsFactors = FALSE)
}
```

@@

```{r}
out <- merge(dat, ldply(out), by="city")
```

@@

```{r}
out <- transform(out, lat = (minlat+maxlat)/2, lon = (minlon+maxlon)/2)
```

>>>>

```{r load}
library("rgbif")
library("ggplot2")
library("plyr")
library("httr")
library("RColorBrewer")
```

@@

Bounding lat/long data is from <https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt>.

```{r box}
url <- 'https://raw.githubusercontent.com/amyxzhang/
boundingbox-cities/master/boundbox.txt'
rawdat <- content(GET(sub("\n", "", url)), as = "text")
dat <- read.table(
  text = rawdat, header = FALSE, 
  sep = "\t", col.names = c("city","minlat","maxlon","maxlat","minlon"),
  stringsAsFactors = FALSE)
dat <- data.frame(
  city = dat$city, minlon = dat$minlon, 
  minlat = dat$minlat, maxlon = dat$maxlon, 
  maxlat = dat$maxlat,
  stringsAsFactors = FALSE
)
```

A helper function to get count data. GBIF has a count API, but we can't use
that with a geometry search as that API doesn't support geospatial search.
We can however use the search API via `occ_search()` and set `limit = 1` so 
that we

```{r newsearch}
getdata <- function(x){
  coords <- as.numeric(x[c('minlon','minlat','maxlon','maxlat')])
  num <- occ_search(geometry = coords, limit = 1)$meta$count
  data.frame(
    city = x['city'], 
    richness = num, 
    stringsAsFactors = FALSE
  )
}
```

@@

```{r}
out <- merge(dat, ldply(out), by = "city")
```

@@

```{r}
out <- transform(out, lat = (minlat + maxlat)/2, lon = (minlon + maxlon)/2)
```

### 2016-08-05 use-cases/niche-modelling.Rmd
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

```{r}
library("rgbif")
library("dismo")
library("maptools")
library("plyr")
```

@@

```{r cache=TRUE}
nn <- name_lookup("bradypus*", rank = "species")
nn <- na.omit(unique(nn$data$nubKey))
df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
df <- df[ sapply(df, function(x) class(x$data)) %in% "data.frame" ]
df <- ldply(lapply(df, "[[", "data"))
df2 <- df[,c('decimalLongitude','decimalLatitude')]
```

>>>>

```{r}
library("rgbif")
library("dismo")
library("maptools")
library("dplyr")
```

@@

```{r cache=TRUE}
nn <- name_lookup("bradypus*", rank = "species")
nn <- na.omit(unique(nn$data$nubKey))
df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
df_data <- df[ sapply(df, function(x) any(class(x$data) %in% "tbl_df")) ]
df_data <- dplyr::bind_rows(lapply(df_data, "[[", "data"))
df2 <- df_data %>% dplyr::select(decimalLongitude, decimalLatitude)
```
